Artificial ecosystem selection reveals relationships between microbiome composition and ecosystem function

### 2022-11-09 Manuscript/manuscript.Rmd
save manuscript progress

Microbiomes mediate important ecosystem functions, yet it has proven difficult
to determine the relationship between microbiome composition and the rate of
ecosystem functions. This challenge remains because it is difficult to
manipulate microbiome composition directly, we often cannot know a priori which
members of a microbiome influence the rate of an ecosystem function, and
microbiomes can covary strongly with other drivers of ecosystem function, such
as the environment. We propose that artificial ecosystem selection can overcome
these challenges. Artificial ecosystem selection involves creating replicate
ecosystems, selecting replicates with a desired ecosystem-level trait, and
using the microbiomes from the selected ecosystems to inoculate a new set of
ecosystems. We performed artificial ecosystem selection on replicate soil
microcosms to select for increased soil methane oxidation rate. We observed a
strong response to selection with a 51% increase in methane oxidation rate per
round over five rounds of selection. This suggests that a large portion of the
variation in methane oxidation rate in these soils can be attributed to
microbiome variation. We also found that selection did not enrich for known
methane oxidizers; rather microbes that responded to selection included members
of `r length(da_fams)` families such as *Fimbriimonadaceae*, *Cytophagaceae*,
and *Diplorickettsiaceae*. This result is in contrast to the typical assumption
that the rate of an ecosystem function is limited by the final step in the
associated microbial pathway. Our study demonstrates that variation in
microbiome composition can contribute to variation in the rate of ecosystem
function and that this relationship can be revealed via artificial ecosystem
selection.

@@

Microbiomes mediate a variety of important functions in ecosystems, and there
is great interest in understanding how attributes of microbiomes may influence
variation in ecosystem functions [@conrad1996; @schimel1998; @crowther2019]. Biodiversity-ecosystem
function relationships have been described for a variety of macroorganismal
communities. For example, plant species richness correlates with productivity
and marine community diversity correlates with functions such as production,
consumption, and nutrient release
[@cardinale2006; @hooper2012; @gamfeldt2015]. While there is some evidence that
variation in microbiomes can contribute to variation in ecosystem functions,
documenting this relationship has been elusive [@conrad1996; @rocca2015; @graham2016].

There are several reasons for this difficulty. One reason is that it is
challenging to manipulate microbiome composition and diversity directly.
Experimental manipulations of community composition and diversity have been
central to documenting the relationships between ecosystem functions and the
biodiversity of plants and animals [@hooper2012], but this is seldom possible
with microbiomes. Most microorganisms cannot yet be cultured and most
available methods for manipulation of microbiomes (e.g., antibiotics, dilution,
etc.) are not sufficiently specific to manipulate individual microbial
lineages. Microbial ecologists are thus often restricted to using comparative
methods, in which microbiome composition and diversity is surveyed via
environmental DNA and then correlated with the rate of an ecosystem function
{cite examples}. However, such studies are rarely successful [@graham2016]. In an
attempt to increase the likelihood of detecting such a correlation, microbial
ecologists may focus their survey efforts on a particular group of microbes
thought to be involved in the ecosystem function of interest. For example, many
studies have looked for relationships between an ecosystem function and the
abundance or diversity of a microbial gene that codes for an enzyme or
enzyme-subunit known to be involved in a particular function. But this approach
assumes that one can identify the “right” marker gene a priori which, given how
little is understood about microbial diversity, seems highly unlikely. And in
fact these studies are rarely successful [@rocca2015].

@@

The problem of connecting microbiome attributes to the rate of an ecosystem
function is analogous to the problem of connecting genomic variation to
phenotypic traits in organisms, and solutions to this problem may be found by
exploring this analogy [@morris2020]. For example, one approach to the problem
of genotype-phenotype mapping is to perform artificial selection on a trait of
interest and identify genes that respond to selection. Such genes are likely to
be causally linked to the trait under selection. An analogous approach for
exploring microbiome-ecosystem function relationships is gaining acceptance
among microbiologists [@swenson2000; @panke-buisse2015; @sanchez2021].
Microbiomes have been shown to respond to selection at the ecosystem-trait
level [@swenson2000], and some ecosystem-scale microbiome traits have been
shown to be transferable through microbiome inoculations [@panke-buisse2015].
Applying artificial selection to whole microbiomes could be a novel approach
for quantifying how much of the variation in the rate of an ecosystem function
could be attributed independently to microbiome variation [@goodnight2000]. In
addition, this approach could be used to identify microbiome attributes, such
as genes or taxa, associated with the rate of an ecosystem function, although,
to our knowledge this has not previously been accomplished [@arias-sanchez2019;
@morris2020]. Such an approach has the potential to expand our ability to more
accurately model microbial ecosystem-function relationships and allow us to
manage ecosystems for particular outcomes [@fuhrman2009; @schimel1998].

@@

We applied artificial ecosystem selection on soil microbiomes by selecting on
the ecosystem-scale CH~4~ oxidation rate.  Using this approach, we were able to
estimate how much of the variation in CH~4~ oxidation rate was associated with
variation in the microbiome.  To identify which microorganisms could be markers
of CH~4~ oxidation rate, we compared the composition of the artificially
selected microbiomes to a control set of microbiomes with random selection.  We
then evaluated whether these markers met our underlying assumptions about
which taxa limit the rate of ecosystem function.

@@

We performed an artificial ecosystem selection experiment [sensu @swenson2000]
by passaging replicate soil microbiomes. The trait we selected on was
CH~4~ oxidation rate. This experiment had two selection lines with twelve jars
each for a total of twenty-four jars per passage. One line was a positive
selection line where the two or three jars with the highest CH~4~ oxidation rate
were chosen to inoculate the next set of positive jars. The other line was a
neutral selection line where an equal number of jars as the positive line were
chosen at random to inoculate the next set of neutral jars. The number of jars
chosen was based on the distribution of fluxes among the positive jars, i.e., we
chose the top three jars unless only two jars had considerably greater CH~4~ oxidation
rates based on visual inspection of histograms. The experiment was carried out
over five passages until a significant divergence in functional rates was 
observed between the two selection treatments.

@@

Bioinformatics processing was performed in 'R' [@rcoreteam2018].
Demultiplexed sequencing reads were denoised using 'DADA2' to generate a table
of amplicon sequence variants (ASVs)
[@callahan2016]. Taxonomic assignment was performed
using the Ribosomal Database Project naive Bayesian classifier [@wang2007]. The presence of
contaminants was evaluated using both the prevalence and frequency methods from
'DECONTAM' [@davis2018].
Decontam identified `r length(which(contamdf$contaminant))` potential contaminants based
on prevalence and frequency. Visual inspection of abundance-concentration
plots indicated that `r length(contaminants)` of these were likely contaminants and these
ASVs were removed. Amplicon sequence variants that were assigned chloroplast taxonomy were removed prior to analysis.

@@

To estimate the proportion of variance in CH~4~ oxidation rate due to variation in
the microbiome, we estimated narrow-sense heritability ($h^2$) as the
regression of offspring on mid-parent [@falconer1996]. The mid-parent was the
mean for all three selected jars and the offspring was the CH~4~ oxidation rate for each jar produced by those parents. First, we tested if there was an effect of treatment
on the heritability estimate. We compared nested models with and without treatment using an F-test
test with the 'anova' function. We then fit the full model to estimate the
heritability of the Positive and Neutral treatments.

@@

We observed a response to selection on whole-ecosystem soil CH~4~ oxidation
rate (`r figures("response", display = "cite")`; difference of slopes:
F~`r compare_response_mods[2,4]`,`r compare_response_mods[2,2]`~ = `r myround(compare_response_mods[2,5], 2)`, p = `r myround(compare_response_mods[2,6], 2)`). At the start of the experiment, the Positive
treatment had a mean CH~4~ oxidation rate that was
`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`\% lower than the
Neutral treatment `r print_lm(response_model, 3, 'difference of y-intercepts')`. 
There was no change in CH~4~ oxidation rate of the Neutral
treatment over the five passages `r print_lm(response_model, 2)`. By contrast,
the Positive treatment had a `r round(per_slope)`\% increase in CH~4~
oxidation rate per passage `r print_lm(response_model, 4)`.

We estimated the proportion of variation in methane oxidation due to variation
in the microbiome as the regression of offspring on mid-parent 
(`r figures("herit", display = "cite")`).  Offspring CH~4~ oxidation rates were
correlated with parental CH~4~ oxidation rates in both the Positive treatment
`r print_lm(herit_model, 4)` and the Neutral treatment `r print_lm(herit_model,
2)`.

@@

While it is well understood that the microbiome mediates a variety of ecosystem
functions, a fundamental question in microbial ecology is whether we need to
understand variation in the composition of microbiomes to understand variation
in the rate of ecosystem functions [@conrad1996; @schimel1998]. To address this question, we performed
artificial ecosystem selection on soil microbiomes by selecting for ecosystems
with high methane oxidation rates. We found that there was a significant
increase in methane oxidation rate as a response to selection. Each generation
of soil ecosystems was created from the same pool of sterile soil substrate and
a small inoculum from the previous generation. Therefore, it is unlikely that
this response to selection was due to changes in the abiotic conditions of the
ecosystem. In addition, the response was only observed in the Positive
treatment and not in the Neutral treatment, which shows that this was not a
side effect of passaging in jars.  Therefore, we conclude that variation in the
microbiome can generate variation in ecosystem functions independent of the
environment.

In quantitative genetics, heritability quantifies the degree to which variation
in a trait is due to variation in genotype [@falconer1996]. This is often estimated as the
regression of offspring phenotype on parental phenotype. Thus, a significant
correlation between the trait values of parents and offspring indicates that
the trait is at least partially genetically determined. To estimate the
variation in an ecosystem trait due to variation in the microbiome, we can
perform a similar estimation by comparing the trait values of the "offspring"
jars in one generation to the trait value of the "parental" jars chosen for
selection in the previous generation. This "ecosystem heritability" then
quantifies how much of the variation in ecosystem function is due to variation
in the microbiome [@goodnight2000; @sanchez2021; @henry2021].

We quantified how much of the variation in methane oxidation
rate was due to variation in the microbiome by regressing the methane oxidation
rates of the 12 offspring jars against the mean methane oxidation rate of the
three jars selected in the previous generation 
We observed a heritability of `r myround(coefficients(herit_model)[4], 2)` (add SE/CI) for the
Positive selection treatment. 
A heritability of 1 would indicate that a trait is entirely
genetically determined and thus offspring would have the exact same value of
that trait as their parents. A heritability greater than 1 indicates that
offspring exceed the value of their parents. This is unlikely in the context of
a genomic trait, but microbiomes are not limited by the constraints of genomes
because microbes are not alleles and communities are not chromosomes.
Therefore, complementarity within a microbial community could result in
offspring communities that exceed the trait values of their "parent"
ecosystems. The large heritability we observed indicates that microbiome
variation is a large contributor to variation in methane oxidation in this
experimental system.

Next, we wanted to address what aspects of the microbiome, in terms of alpha
diversity, beta diversity, or individual taxa, explain this contribution of the
microbiome to variation in methane oxidation rate. There are three ways that
the microbiome could respond to selection in this experiment: gain or loss of
species, changes in the relative abundances of species, or changes within the
genomes of the constituent species. Using 16S rRNA gene sequences, we are
unable to address the third possibility of whether taxa in this experiment
could have evolved changes in their genomes as a result of selection.
Therefore, we will focus on the first two possibilities.

@@

Even though we observed an increase in methane oxidation rate in the Positive
treatment and a difference in composition between the two treatments, we did
not observe an increase in the relative abundance of methanotrophs. This was
surprising given that methane consumption is not a common trait among microbes
and that it is often assumed that the rate of an ecosystem function is
limited by the final enzymatic step in a metabolic pathway
[@rocca2015]. In certain ecosystems, methane production and consumption are 
correlated with the abundance of methanogens and methanotrophs as estimated from marker
genes [@meyer2020, @meyer2020a]. However, it does not appear generally true that the rate of an
ecosystem function is limited by the abundance of the gene that encodes the
final step in that pathway [@rocca2015]. Our results suggest that the rate of methane oxidation can be altered by
non-methanotrophs, perhaps through ecological interactions with methanotrophic
species.

@@

As more is understood about the role of the microbiome in regulating ecosystem
functions and the specific taxa involved, more detailed experiments could
selectively add or remove important taxa or manipulate their relative abundance
to start to estimate the quantitative relationship between specific taxa and
the rate of an ecosystem function. Both our differential abundance analysis and
this reductionist approach to manipulating individual taxa assumes that taxa
act individually on ecosystem function. However, it is much more likely that
the microbiome affects the rate of ecosystem function through the interactions
of multiple taxa. We are unable to evaluate this possibility given our small 
sample size, but future analyses with greater replication should test the
role of multiple taxa simultaneously by, for example, using machine learning
methods [@topcuoglu2020].

A fundamental question in microbial ecology is whether variation in the
microbiome contributes to variation in ecosystem function. We performed
artificial ecosystem selection on soil methane oxidation rate to test the
relationship between microbiome variation and soil methane oxidation rate. We
observed a response to selection on soil methane oxidation rate and found a
high heritability for methane oxidation, which suggests that variation in the
microbiome contributes to variation in soil methane oxidation rate.
Surprisingly, we did not observe an increase in the relative of abundance of
methanotrophs as a response to selection. This suggests that the rate of methane 
oxidation could be altered by
non-methanotrophs. Future research could estimate
heritability under greater environmental variation to understand how much of
the variation in methane oxidation rate is due to variation in the microbiome
in a natural ecosystem. In addition, future selection experiments could use
genome-resolved metagenomics to observe within-strain evolution and test
changes in the functional pathways of the microbiome as a response to selection
to determine what aspects of the microbiome explain this increase in oxidation
rate. This study demonstrates that soil methane oxidation rate can vary greatly
with variation in the microbiome and opens up future research opportunities to
establish a more detailed understanding of the relationship between the
microbiome and ecosystem function.

@@

`r figures("response", "Response to selection on soil CH~4~ oxidation rate fit by ordinary least squares regression. The y-axis is the log~10~ of the additive inverse of the first-order exponential decay constant $k$ (i.e., $-k$) with units day^-1^. The x-axis is the passage number. The orange line and points are the Positive treatment and the gray line and points are the Neutral treatment.")`

@@

p_herit_each

@@

`r figures("herit", "Ordinary least squares regression of offspring CH~4~ flux on mid-parent CH~4~ flux ($-k$ with units day^-1^). Mean parental is the mean of the jars selected to inoculate the next passage. Offspring is the oxidation rate of each jar in the next passage. The orange lines and points are the Positive treatment and the gray lines and points are the Neutral treatment")`

>>>>

Microbiomes mediate important ecosystem functions, yet it has proven difficult
to determine the relationship between microbiome composition and the rate of
ecosystem functions. This challenge remains because it is difficult to
manipulate microbiome composition directly, we often cannot know a priori which
members of a microbiome influence the rate of an ecosystem function, and
microbiomes can covary strongly with other drivers of ecosystem function, such
as the environment. We propose that artificial ecosystem selection can overcome
these challenges. Artificial ecosystem selection involves creating replicate
ecosystems, selecting replicates with a desired ecosystem-level trait, and
using the microbiomes from the selected ecosystems to inoculate a new set of
ecosystems. We performed artificial ecosystem selection on replicate soil
microbiomes to select for increased soil methane oxidation rate. We observed a
strong response to selection with a `r myround(per_slope, 1)`\% increase in
methane oxidation rate per round over five rounds of selection. In addition, we
estimated that `r myround((coefficients(herit_model)[2] +
coefficients(herit_model)[4])*100, 1)`\% of the variation in methane oxidation
rate in these soils can be attributed to microbiome variation. We also found
that selection did not enrich for known methane oxidizers; rather microbes that
responded to selection included members of `r length(da_fams)` families not
known to oxidize methane, such as
*Fimbriimonadaceae*, *Cytophagaceae*, and *Diplorickettsiaceae*. This result is
in contrast to the typical assumption that the rate of an ecosystem function is
limited by the final step in the associated microbial pathway. Our study
demonstrates that variation in microbiome composition can contribute to
variation in the rate of ecosystem function and that manipulating microbiome composition
could be a viable strategy for managing ecosystem functions.

@@

Microbiomes mediate a variety of important functions in ecosystems, and there
is great interest in understanding how attributes of microbiomes may influence
variation in ecosystem functions [@conrad1996; @schimel1998; @crowther2019].
Biodiversity-ecosystem function relationships have been documented for a
variety of macroorganismal communities. For example, plant species richness
correlates with productivity and marine community diversity correlates with
functions such as production, consumption, and nutrient release
[@cardinale2006; @hooper2012; @gamfeldt2015]. While there is some evidence that
variation in microbiomes can contribute to variation in ecosystem functions,
describing a more detailed mapping between the microbiome and ecosystem
function has been elusive [@conrad1996; @rocca2015; @graham2016].

There are several reasons for this difficulty. One reason is that it is
challenging to manipulate microbiome composition and diversity directly.
Experimental manipulations of community composition and diversity have been
central to documenting the relationships between ecosystem functions and the
biodiversity of plants and animals [@hooper2012], but this is seldom possible
with microbiomes. Most microorganisms cannot yet be cultured and most
available methods for manipulation of microbiomes (e.g., antibiotics, dilution,
etc.) are not sufficiently specific to manipulate individual microbial
lineages. Microbial ecologists are thus often restricted to using comparative
methods, in which microbiome composition and diversity is surveyed via
environmental DNA and then correlated with the rate of an ecosystem function. 
However, microbiome attributes add an average of only 8% of variance explained
to ecosystem models [@graham2016].

In an
attempt to increase the likelihood of detecting such a correlation, microbial
ecologists may focus their survey efforts on a particular group of microbes
thought to be involved in the ecosystem function of interest. For example, many
studies have looked for relationships between an ecosystem function and the
abundance or diversity of a microbial gene that codes for an enzyme or
enzyme-subunit known to be involved in a particular function. But this approach
assumes that one can identify the “right” marker gene a priori which, given how
little is understood about microbial diversity, seems highly unlikely. And for
most functions in most ecosystems there is no correlation between the abundance
of a marker gene and the rate of the corresponding process [@rocca2015].

@@

The problem of connecting microbiome attributes to the rate of an ecosystem
function is analogous to the problem of connecting genomic variation to
phenotypic traits in organisms, and solutions to this problem may be found by
exploring this analogy [@morris2020]. For example, one approach to the problem
of genotype-phenotype mapping is to perform artificial selection on a trait of
interest and identify genes that respond to selection. Such genes are likely to
be causally linked to the trait under selection. An analogous approach for
exploring microbiome-ecosystem function relationships is gaining acceptance
among microbiologists [@swenson2000; @blouin2015; @panke-buisse2015; @sanchez2021].
Microbiomes have been shown to respond to selection at the ecosystem-trait
level [@swenson2000], and some ecosystem-scale microbiome traits have been
shown to be transferable through microbiome inoculations [@panke-buisse2015].
Applying artificial selection to whole microbiomes could be a novel approach
for quantifying how much of the variation in the rate of an ecosystem function
could be attributed independently to microbiome variation [@goodnight2000; @blouin2015]. 
By "microbiome variation," we mean aspects of microbial diversity that are
routinely measured by microbial ecologists, such as the relative
abundance of taxonomic groups inferred from amplicon sequence variants. 
In
addition, this approach could be used to identify microbiome attributes, such
as genes or taxa, associated with the rate of an ecosystem function, although,
to our knowledge this has not previously been accomplished [@arias-sanchez2019].
Such an approach has the potential to expand our ability to more
accurately model microbial ecosystem-function relationships and allow us to
manage ecosystems for particular outcomes [@fuhrman2009; @schimel1998].

@@

In this study, we wanted to evaluate whether variation in the microbiome
contributes to variation in the rate of ecosystem functions and to identify
microbiome attributes that might explain the mapping between microbiome and
ecosystem function. To begin to address this problem, we used artificial
ecosystem selection on soil microbiomes as a model system.  Using this
experimental model, we answered the following questions: Does variation in the
relative abundance of microbial taxa contribute to variation in soil methane
oxidation rate in the lab? How much of the variation in soil methane oxidation
rate can be attributed to the microbiome? What attributes of the microbiome
regulate variation in methane oxidation rate? And do these attributes match our
assumptions about what regulates methane oxidation rate?  We applied artificial
ecosystem selection on soil microbiomes by selecting on the ecosystem-scale
CH~4~ oxidation rate. We estimated how much of the variation in CH~4~ oxidation
rate was associated with variation in the microbiome by calculating the slope
of the regression between donor and recipient communities. To identify which
microorganisms could be markers of CH~4~ oxidation rate, we compared the
composition of the artificially selected microbiomes to a control set of
microbiomes with random selection. We then evaluated whether these markers met
our underlying assumptions about which taxa limit the rate of ecosystem
function.

@@

We performed an artificial ecosystem selection experiment [*sensu* @swenson2000]
by passaging replicate soil microbiomes. The trait we selected on was
CH~4~ oxidation rate. This experiment had two selection lines with twelve jars
each for a total of twenty-four jars per passage. One line was a positive
selection line where the two or three jars with the highest CH~4~ oxidation rate
were chosen to inoculate the next set of positive jars. The other line was a
neutral selection line where an equal number of jars as the positive line were
chosen at random to inoculate the next set of neutral jars. The number of jars
chosen was based on the distribution of fluxes among the positive jars, i.e., we
chose the top three jars unless only two jars had considerably greater CH~4~ oxidation
rates based on visual inspection of histograms. The experiment was carried out
over five passages until a significant divergence in functional rates was 
observed between the two selection treatments.

@@

Bioinformatics processing was performed in 'R' [@rcoreteam2018].
Demultiplexed sequencing reads were denoised using 'DADA2' to generate a table
of amplicon sequence variants (ASVs)
[@callahan2016]. Taxonomic assignment was performed
using the Ribosomal Database Project naive Bayesian classifier [@wang2007]. The presence of
contaminants was evaluated using both the prevalence and frequency methods from
'DECONTAM' [@davis2018].
Decontam identified `r length(which(contamdf$contaminant))` potential contaminants based
on prevalence and frequency. Visual inspection of abundance-concentration
plots indicated that `r length(contaminants)` of these were likely contaminants and these
ASVs were removed. Amplicon sequence variants that were assigned chloroplast or mitochondria taxonomy were removed prior to analysis.

@@

We estimated the proportion of variance in CH~4~ oxidation rate due to variation in
the microbiome as the
regression of recipient on donor [@falconer1996]. The mid-donor was the
mean for all three selected jars and the mid-recipient was the mean methane oxidation 
rate for the twelve jars produced by those three donors. First, we tested if there was an effect of treatment by comparing nested models with and without treatment using an F-test with the 'anova' function. We then fit the full model to estimate the
slope of the relationship between recipient and donor for the Positive and Neutral treatments.

@@

We observed a response to selection on whole-ecosystem soil CH~4~ oxidation
rate (`r figures("response", display = "cite")`; difference of slopes:
F~`r compare_response_mods[2,4]`,`r compare_response_mods[2,2]`~ = `r myround(compare_response_mods[2,5], 2)`, p = `r myround(compare_response_mods[2,6], 2)`). At the start of the experiment, the Positive
treatment had a mean CH~4~ oxidation rate that was
`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`\% lower than the
Neutral treatment `r print_lm(response_model, 3, 'difference of y-intercepts')`. 
There was no change in CH~4~ oxidation rate of the Neutral
treatment over the five passages `r print_lm(response_model, 2)`. By contrast,
the Positive treatment had a `r myround(per_slope, 1)`\% increase in CH~4~
oxidation rate per passage `r print_lm(response_model, 4)`.

We estimated the proportion of variation in methane oxidation due to variation
in the microbiome as the regression of mid-recipient on mid-donor 
(`r figures("herit", display = "cite")`).  Recipient CH~4~ oxidation rates were
correlated with donor CH~4~ oxidation rates in both the Positive treatment
`r paste0('(', 'slope', ' = ', myround(tidy(herit_model)[2, 2] + tidy(herit_model)[4, 2], 2), ', SE = ', myround(tidy(herit_model)[4, 3], 2), ', t = ', myround(tidy(herit_model)[4, 4], 2), ', p = ', myround(tidy(herit_model)[4, 5], 2), ')')` and 
the Neutral treatment `r print_lm(herit_model, 2)`.

@@

While it is well understood that the microbiome mediates a variety of ecosystem
functions, a fundamental question in microbial ecology is whether we need to
understand variation in the composition of microbiomes to understand variation
in the rate of ecosystem functions [@conrad1996; @schimel1998]. To address this
question, we performed artificial ecosystem selection on soil microbiomes by
selecting for ecosystems with high methane oxidation rates. We found that there
was a significant increase in methane oxidation rate in the Positive selection
treatment relative to the Neutral selection treatment. This suggests that 
variation in the microbiome can generate variation in ecosystem
functions independent of the environment.

We wanted to estimate how much of the variation in ecosystem function can be
attributed to variation in the microbiome. This problem is analogous to the
problem in quantitative genetics of estimating how much of the variation in an
organismal trait is due variation in genotype, i.e., heritability. Organismal
biologists often estimate the percent heritability of a trait as the slope of the
regression of offspring phenotype
on parental phenotype [@falconer1996]. A correlation between the trait values 
of parents and offspring indicates that the trait is at least partially 
genetically determined. Using artificial
ecosystem selection we can estimate the variation in an ecosystem trait due to
variation in the microbiome by comparing the trait values of the "offspring" or
recipient jars in one generation to the trait value of the "parental" or donor
jars chosen for selection in the previous generation. This "ecosystem
heritability" then quantifies what percentage of the variation in ecosystem function
is due to variation in the microbiome [@goodnight2000; @blouin2015].

We quantified how much of the variation in methane oxidation rate was due to
variation in the microbiome by regressing the mean methane oxidation rates of
the 12 recipient jars against the mean methane oxidation rate of the three
donor jars selected to inoculate the recipients. We observed a heritability of
`r myround((coefficients(herit_model)[2] + coefficients(herit_model)[4])*100,
1)`\% (SE = `r myround(tidy(herit_model)$std.error[4] * 100, 1)`\%) for the
Positive selection treatment. This estimate is comparable to the heritability
of carbon dioxide emissions from water [@blouin2015]. Notably, we observed a
negative heritability in the Neutral treatment 
(`r myround(coefficients(herit_model)[2]*100, 1)`\%, 
SE = `r myround(tidy(herit_model)$std.error[2] * 100, 1)`\%). Negative heritability
estimates are not uncommon in quantitative genetics [@steinsaltz2020]. This can
occur for random selection controls in artificial selection experiments and may
indicate that methane oxidation rate is negatively correlated with other traits
associated with persistence in jars. 
The heritability we observed in the Positive treatment indicates that
roughly half of the variation in methane oxidation rate is due to variation in
the microbiome in our experimental system.

Given that variation in the microbiome is driving the observed variation in
methane oxidation rate, we next wanted to address what aspects of the microbiome, in
terms of alpha diversity, beta diversity, or the relative abundance of taxa,
explain this contribution of the
microbiome to variation in methane oxidation rate. There are three ways that
the microbiome could respond to selection in this experiment: gain or loss of
species, changes in the relative abundances of species, or changes within the
genomes of the constituent species. Using 16S rRNA gene sequences, we are
unable to address the third possibility of whether taxa in this experiment
could have evolved changes in their genomes as a result of selection.
Therefore, we will focus on the first two possibilities.

@@

Even though we observed an increase in methane oxidation rate in the Positive
treatment and a difference in composition between the two treatments, we did
not observe an increase in the relative abundance of methanotrophs. This was
surprising given that methane consumption is not a common trait among microbes
and that it is often assumed that the rate of an ecosystem function is limited
by the final enzymatic step in a metabolic pathway [@rocca2015]. In certain
ecosystems, methane production and consumption are correlated with the
abundance of methanogens and methanotrophs as estimated from marker genes
[@meyer2020; @meyer2020a]. However, it does not appear generally true that the
rate of an ecosystem function is limited by the abundance of the gene that
encodes the final step in that pathway [@rocca2015]. Our results suggest that
ecosystem-scale methane oxidation rates can be altered by non-methanotrophs, perhaps
through ecological interactions with methanotrophic species.

@@

A fundamental question in microbial ecology is whether variation in the
microbiome contributes to variation in ecosystem function. We performed
artificial ecosystem selection on soil methane oxidation rate to test the
relationship between microbiome variation and soil methane oxidation rate. We
observed a response to selection on soil methane oxidation rate and found a
high heritability for methane oxidation, which suggests that variation in the
microbiome contributes to variation in soil methane oxidation rate
independent of the environment. This shows that manipulating the microbiome
could be a viable strategy for mitigating or enhancing certain
ecosystem functions.  Surprisingly, we did not observe an increase in the
relative abundance of methanotrophs as a response to selection. This
suggests that the rate of methane oxidation could be altered by
non-methanotrophs, possibly through ecological interactions with methanotrophs.
Future research should investigate other aspects of microbiome variation, such
as strain-level evolution or functional gene composition, to begin to describe
a more detailed mapping between the microbiome and ecosystem function. This
study demonstrates that soil methane oxidation rate can vary greatly with
variation in the microbiome and opens up future research opportunities to
establish a more detailed understanding of the relationship between the
microbiome and ecosystem function.

@@

`r figures("response", "Response to selection on soil CH~4~ oxidation rate. The y-axis is the log~10~ of the additive inverse of the first-order exponential decay constant $k$ (i.e., $-k$) with units day^-1^. The x-axis is the passage number. The orange line and points are the Positive treatment and the gray line and points are the Neutral treatment.")`

@@

p_herit

@@

`r figures("herit", paste0("Regression of mid-recipient methane flux on mid-donor methane flux ($-k$ with units day^-1^). Donor is the mean oxidation rate of the jars selected to inoculate the next passage. Recipient is the mean oxidation rate of the twelve jars in the next passage. The orange lines and points are the Positive treatment and the gray lines and points are the Neutral treatment. Slope of the regression line estimates the proportion of variation in methane oxidation rate due to variation in the microbiome. For the Positive treatment, the variance explained by the microbiome was ", myround((coefficients(herit_model)[2] + coefficients(herit_model)[4])*100, 1), "% (SE = ", myround(tidy(herit_model)$std.error[4] * 100, 1), "%) and for the Neutral treatment the variance explained by the microbiome was ", myround(coefficients(herit_model)[2]*100, 1), "% (SE = ", myround(tidy(herit_model)$std.error[2] * 100, 1), "%)."))`
